---
title: "Checking_for_deletions"
author: "Hugo Barreto"
date: "2023-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rstudioapi)
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
print( getwd() )

if(!require(librarian)){
    install.packages("librarian")
    library(librarian)
}

librarian::shelf(dplyr, stringr, tidyverse, XML, RCurl, rlist, janitor, writexl, ggbeeswarm, Rsamtools, gmoviz,ggpubr)

if("plyr" %in% (.packages())){ # Due to conflicts with plyr, this code deactivates this package
  detach("package:plyr", unload=TRUE) 
}

```

## Some mice have the indication of possible deletions

### This script calculates the frequency of the deletions

### Calculate the frequency of the deletion [PROKKA_00047]–[PROKKA_00125] deletion (position 47576 to 134881, Δ87,305 bp) in all mice


```{r [PROKKA_00047]–[PROKKA00125], warning=FALSE}

list_of_populations_bam_I <- dir(pattern = "*pop_R.bam$",full.names = T)

coverage_total_setC_PROKKA_00125 <- data.frame()

# Get coverage from the region of the deletion, including flanking genes
for (population in list_of_populations_bam_I) {
  
  setC_PROKKA_00125_region<-GRanges('NC_000913', IRanges(45731, 136627))

  setC_PROKKA_00125_coverage<-as.data.frame(getCoverage(regions_of_interest = setC_PROKKA_00125_region,bam_file = population,window_size = 1))

  setC_PROKKA_00125_coverage <- setC_PROKKA_00125_coverage %>% 
  mutate(population = population) %>%
  dplyr::rename(position = start)

setC_PROKKA_00125_coverage$population <- gsub("_pop_R.bam","",setC_PROKKA_00125_coverage$population) #remove .html from the cells
setC_PROKKA_00125_coverage$population <- gsub("./","",setC_PROKKA_00125_coverage$population) #remove input/ from the cells

coverage_total_setC_PROKKA_00125 <- rbind(coverage_total_setC_PROKKA_00125,setC_PROKKA_00125_coverage)
  
}

coverage_total_setC_PROKKA_00125[c('mouse', 'day')] <- str_split_fixed(coverage_total_setC_PROKKA_00125$population, 'd', 2)
coverage_total_setC_PROKKA_00125 <- coverage_total_setC_PROKKA_00125 %>% 
  relocate(mouse,day,.before = seqnames) %>%
  dplyr::select(-population,-seqnames,-end,-width,-strand)  
  
coverage_total_setC_PROKKA_00125$gene <- cut(coverage_total_setC_PROKKA_00125$position,
                             breaks = c(45731,47575,
                                        59330,60901,
                                        61027,61239,
                                        61556,64067,
                                        66118,67375,
                                        71114,82720,
                                        86891,92333,
                                        96326,100921,
                                        104266,105523,
                                        108588,109973,
                                        111007,114457,
                                        
                                        
                                        135245,136627),
                             labels = c("setC","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion","yicJ"
                                        ))

coverage_total_setC_PROKKA_00125 <- coverage_total_setC_PROKKA_00125[!(is.na(coverage_total_setC_PROKKA_00125$gene) | coverage_total_setC_PROKKA_00125$gene==""), ]


setC_PROKKA_00125_medians <- coverage_total_setC_PROKKA_00125 %>% group_by(mouse,day,gene) %>%
  summarize(median=median(coverage))


setC_PROKKA_00125_medians <- setC_PROKKA_00125_medians %>% 
  tidyr::pivot_wider(id_cols = c("mouse","day"),
              names_from = "gene",
              values_from = "median")

setC_PROKKA_00125_medians <- setC_PROKKA_00125_medians %>% 
  mutate(excs1 = 1-(deletion/setC), 
         excs2 = 1-(deletion/yicJ), 
         position = 47576,
         mutation = "Δ87,305 bp",
         annotation = "deletion",
         freq = ifelse(mean(c(excs1,excs2)) < 0, 0, mean(c(excs1,excs2))),
         gene = "[PROKKA_00047]–[PROKKA_00125]",
         day = as.numeric(day),
         description = "79 genes [PROKKA_00047],PROKKA_00048,cbtA_1,cbeA_1,PROKKA_00051,PROKKA_00052,kicA_1,PROKKA_00054,flu_1,era_1,PROKKA_00057,PROKKA_00058,PROKKA_00059,PROKKA_00060,PROKKA_00061,PROKKA_00062,PROKKA_00063,PROKKA_00064,PROKKA_00065,PROKKA_00066,PROKKA_00067,PROKKA_00068,PROKKA_00069,PROKKA_00070,PROKKA_00071,PROKKA_00072,PROKKA_00073,PROKKA_00074,shlB_1,shlA_1,PROKKA_00077,PROKKA_00078,PROKKA_00079,PROKKA_00080,PROKKA_00081,PROKKA_00082,PROKKA_00083,PROKKA_00084,PROKKA_00085,PROKKA_00086,PROKKA_00087,PROKKA_00088,fhaB_1,PROKKA_00090,PROKKA_00091,PROKKA_00092,PROKKA_00093,fhaB_2,PROKKA_00095,PROKKA_00096,PROKKA_00097,PROKKA_00098,PROKKA_00099,PROKKA_00100,PROKKA_00101,PROKKA_00102,PROKKA_00103,PROKKA_00104,PROKKA_00105,PROKKA_00106,PROKKA_00107,PROKKA_00108,PROKKA_00109,PROKKA_00110,PROKKA_00111,yadA_1,PROKKA_00113,PROKKA_00114,icsA_1,PROKKA_00116,PROKKA_00117,gadX_1,PROKKA_00119,PROKKA_00120,virF_1,PROKKA_00112,intS_1,intS_2,[PROKKA_00125]")

setC_PROKKA_00125_medians <- setC_PROKKA_00125_medians %>% dplyr::select(-c(setC,deletion,yicJ,excs1,excs2))

ggplot(coverage_total_setC_PROKKA_00125,aes(x=gene,y=coverage))+
  ggbeeswarm::geom_quasirandom()+
  geom_boxplot(alpha=0.3)+
  cowplot::theme_cowplot()+
  facet_wrap(~mouse+day)

setC_PROKKA_00125_medians

final_setC_PROKKA_00125_medians <- setC_PROKKA_00125_medians %>% filter(freq != 0)

write_xlsx(as.data.frame(final_setC_PROKKA_00125_medians),
              "PROKKA_00047-PROKKA00125_deletion_freq.xlsx")

```


### Calculate the frequency of the deletion [PROKKA_01034]–[PROKKA01105] deletion (position 1081423 to 1167367, Δ85,944 bp) in all mice


```{r [PROKKA_01034]–[PROKKA01105], warning=FALSE}

list_of_populations_bam_I <- dir(pattern = "*pop_R.bam$",full.names = T)

coverage_total_PROKKA_01034_PROKKA01105 <- data.frame()

# Get coverage from the region of the deletion, including flanking genes
for (population in list_of_populations_bam_I) {
  
  PROKKA_01034_PROKKA01105_region<-GRanges('NC_000913', IRanges(1080540, 1168074))

  PROKKA_01034_PROKKA01105_coverage<-as.data.frame(getCoverage(regions_of_interest = PROKKA_01034_PROKKA01105_region,bam_file = population,window_size = 1))

  PROKKA_01034_PROKKA01105_coverage <- PROKKA_01034_PROKKA01105_coverage %>% 
  mutate(population = population) %>%
  dplyr::rename(position = start)

PROKKA_01034_PROKKA01105_coverage$population <- gsub("_pop_R.bam","",PROKKA_01034_PROKKA01105_coverage$population) #remove .html from the cells
PROKKA_01034_PROKKA01105_coverage$population <- gsub("./","",PROKKA_01034_PROKKA01105_coverage$population) #remove input/ from the cells

coverage_total_PROKKA_01034_PROKKA01105 <- rbind(coverage_total_PROKKA_01034_PROKKA01105,PROKKA_01034_PROKKA01105_coverage)
  
}

coverage_total_PROKKA_01034_PROKKA01105[c('mouse', 'day')] <- str_split_fixed(coverage_total_PROKKA_01034_PROKKA01105$population, 'd', 2)
coverage_total_PROKKA_01034_PROKKA01105 <- coverage_total_PROKKA_01034_PROKKA01105 %>% 
  relocate(mouse,day,.before = seqnames) %>%
  dplyr::select(-population,-seqnames,-end,-width,-strand)  
  
coverage_total_PROKKA_01034_PROKKA01105$gene <- cut(coverage_total_PROKKA_01034_PROKKA01105$position,
                             breaks = c(1080540,1081247,
                                        1081423,1081626,
                                        1085772,1087745,
                                        1090483,1112075,
                                        1117297,1117842,
                                        1129115,1147783,
                                        1151144,1154508,
                                        1158986,1162059,
                                        1164639,1167367,
                                        1167538,1168074),
                             
                             labels = c("ydfK",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","epsM"
                                        ))

coverage_total_PROKKA_01034_PROKKA01105 <- coverage_total_PROKKA_01034_PROKKA01105[!(is.na(coverage_total_PROKKA_01034_PROKKA01105$gene) | coverage_total_PROKKA_01034_PROKKA01105$gene==""), ]


PROKKA_01034_PROKKA01105_medians <- coverage_total_PROKKA_01034_PROKKA01105 %>% group_by(mouse,day,gene) %>%
  summarize(median=median(coverage))


PROKKA_01034_PROKKA01105_medians <- PROKKA_01034_PROKKA01105_medians %>% 
  tidyr::pivot_wider(id_cols = c("mouse","day"),
              names_from = "gene",
              values_from = "median")

PROKKA_01034_PROKKA01105_medians <- PROKKA_01034_PROKKA01105_medians %>% 
  mutate(excs1 = 1-(deletion/ydfK), 
         excs2 = 1-(deletion/epsM), 
         position = 1081423,
         mutation = "Δ85,944 bp",
         annotation = "deletion",
         freq = ifelse(mean(c(excs1,excs2)) < 0, 0, mean(c(excs1,excs2))),
         gene = "[PROKKA_01034]–[PROKKA_01105]",
         day = as.numeric(day),
         description = "71 genes [PROKKA_01034],,[PROKKA_01105]")

PROKKA_01034_PROKKA01105_medians <- PROKKA_01034_PROKKA01105_medians %>% dplyr::select(-c(ydfK,deletion,epsM,excs1,excs2))

ggplot(coverage_total_PROKKA_01034_PROKKA01105,aes(x=gene,y=coverage))+
  ggbeeswarm::geom_quasirandom()+
  geom_boxplot(alpha=0.3)+
  cowplot::theme_cowplot()+
  facet_wrap(~mouse+day)

PROKKA_01034_PROKKA01105_medians

final_PROKKA_01034_PROKKA01105_medians <- PROKKA_01034_PROKKA01105_medians %>% filter(freq != 0)

write_xlsx(as.data.frame(final_PROKKA_01034_PROKKA01105_medians),
              "PROKKA_01034-PROKKA01105_deletion_freq.xlsx")

```


### Calculate the frequency of the deletion [PROKKA_01215]–[intA_1] deletion (position 1285432 to 1318675, Δ33,243 bp) in all mice


```{r [PROKKA_01215]–[intA_1], warning=FALSE}

list_of_populations_bam_I <- dir(pattern = "*pop_R.bam$",full.names = T)

coverage_total_PROKKA_01215_intA_1 <- data.frame()

# Get coverage from the region of the deletion, including flanking genes
for (population in list_of_populations_bam_I) {
  
  PROKKA_01215_intA_1_region<-GRanges('NC_000913', IRanges(1284673, 1328920))

  PROKKA_01215_intA_1_coverage<-as.data.frame(getCoverage(regions_of_interest = PROKKA_01215_intA_1_region,bam_file = population,window_size = 1))

  PROKKA_01215_intA_1_coverage <- PROKKA_01215_intA_1_coverage %>% 
  mutate(population = population) %>%
  dplyr::rename(position = start)

PROKKA_01215_intA_1_coverage$population <- gsub("_pop_R.bam","",PROKKA_01215_intA_1_coverage$population) #remove .html from the cells
PROKKA_01215_intA_1_coverage$population <- gsub("./","",PROKKA_01215_intA_1_coverage$population) #remove input/ from the cells

coverage_total_PROKKA_01215_intA_1 <- rbind(coverage_total_PROKKA_01215_intA_1,PROKKA_01215_intA_1_coverage)
  
}

coverage_total_PROKKA_01215_intA_1[c('mouse', 'day')] <- str_split_fixed(coverage_total_PROKKA_01215_intA_1$population, 'd', 2)
coverage_total_PROKKA_01215_intA_1 <- coverage_total_PROKKA_01215_intA_1 %>% 
  relocate(mouse,day,.before = seqnames) %>%
  dplyr::select(-population,-seqnames,-end,-width,-strand)  
  
coverage_total_PROKKA_01215_intA_1$gene <- cut(coverage_total_PROKKA_01215_intA_1$position,
                             breaks = c(1284673,1285258,
                                        1285558,1288017,
                                        1288467,1290591,
                                        1297328,1308598,
                                        1328145,1328369),
                             
                             labels = c("mug",
                                        "","deletion",
                                        "","deletion",
                                        "","deletion",
                                        "","traC_1"
                                        ))

coverage_total_PROKKA_01215_intA_1 <- coverage_total_PROKKA_01215_intA_1[!(is.na(coverage_total_PROKKA_01215_intA_1$gene) | coverage_total_PROKKA_01215_intA_1$gene==""), ]


PROKKA_01215_intA_1_medians <- coverage_total_PROKKA_01215_intA_1 %>% group_by(mouse,day,gene) %>%
  summarize(median=median(coverage))


PROKKA_01215_intA_1_medians <- PROKKA_01215_intA_1_medians %>% 
  tidyr::pivot_wider(id_cols = c("mouse","day"),
              names_from = "gene",
              values_from = "median")

PROKKA_01215_intA_1_medians <- PROKKA_01215_intA_1_medians %>% 
  mutate(excs1 = 1-(deletion/mug), 
         excs2 = 1-(deletion/traC_1), 
         position = 1285432,
         mutation = "Δ33,243 bp",
         annotation = "deletion",
         freq = ifelse(mean(c(excs1,excs2)) < 0, 0, mean(c(excs1,excs2))),
         gene = "[PROKKA_01215]–[intA_1]",
         day = as.numeric(day),
         description = "28 genes [PROKKA_01215],...,[intA_1]")

PROKKA_01215_intA_1_medians <- PROKKA_01215_intA_1_medians %>% dplyr::select(-c(mug,deletion,traC_1,excs1,excs2))

ggplot(coverage_total_PROKKA_01215_intA_1,aes(x=gene,y=coverage))+
  ggbeeswarm::geom_quasirandom()+
  geom_boxplot(alpha=0.3)+
  cowplot::theme_cowplot()+
  facet_wrap(~mouse+day)

PROKKA_01215_intA_1_medians

final_PROKKA_01215_intA_1_medians <- PROKKA_01215_intA_1_medians %>% filter(freq != 0)

write_xlsx(as.data.frame(final_PROKKA_01215_intA_1_medians),
              "PROKKA_01215-intA_1_deletion_freq.xlsx")

```
